---
title: "Generate Figs and Tables"
author: "Lauren White (lauren.white@cdph.ca.gov)"
date: '2024-12-05'
output:
  html_document:
    toc: true
    toc_float: true
---

Code to reproduce Fig and table results in manuscript. Can be run sequentially after `load_data.R`, `calculate_forecastability.Rmd`, and `forecast_eval.Rmd` or as a standalone using output already saved locally to the `/results` folder for convenience.

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggrepel)
library(dplyr)
library(MetBrewer)
library(readr)
library(cowplot)
library(tidyr)
library(purrr)
library(segmented)
library(astsa)
library(ForeCA)
library(forecast)

set.seed(1234)
`%notin%` <- Negate(`%in%`)

season22_start_MMWR<- as.Date("2022-10-02")  #Beginning of week 40 for 2022-2023 season
season23_start_MMWR<- as.Date("2023-10-01") #Beginning of week 40 for 2023-2024 season
season22_start<-  as.Date("2022-07-01")
season23_start<- as.Date("2023-07-01")
season24_end<- as.Date("2024-04-30") #End of mandatory NHSN reporting period in 2024
```

### Settings for `foreCA` `Omega` function
```{r}
# spectrum control
sc <- list(method = "mvspec")

# entropy control
ec <- list(prior.weight = 1e-2)

confidence<-0.05
```

### Load forecastability data frame
(if not already loaded in environment from running other Rmd scripts)
```{r}
if(!exists("forecastability")) {
  forecastability<-read_csv("results/forecastability_results_MMWR.csv")
}

if(!exists("forecastability_horizon")) {
  forecastability_horizon<-read_csv("results/forecastability_horizon_results_MMWR.csv")
}

if(!exists("forecastability_summer")) {
  forecastability_summer<-read_csv("results/forecastability_results.csv")
}

if(!exists("combo")) {
  combo<-read_csv("results/combo_omega.csv")
}

if(!exists("hcai_combined")) {
  hcai_combined<-read_csv("results/hcai_combined.csv")
}

if(!exists("hcai_state_omega")) {
  hcai_state_omega<-read_csv("results/hcai_state_omega.csv")
}

if(!exists("hcai_burden")) {
  hcai_burden<-read_csv("results/hcai_burden.csv")
}

if(!exists("state_covid_cadence")) {
  state_covid_cadence<-read_csv("results/state_covid_cadence.csv")
}

if(!exists("state_flu_cadence")) {
  state_flu_cadence<-read_csv("results/state_flu_cadence.csv")
}

if(!exists("flu_subtypes")) {
flu_subtypes<-read_csv("data/CDPH_flu_subtypes.csv") %>% rename(season=Season)
}

if(!exists("covid_hosp_full")) {
covid_hosp_full<-read_csv("data/covid_hosp.csv") 
}

if(!exists("flu_hosp_full")) {
flu_hosp_full<-read_csv("data/flu_hosp.csv") 
}

if(!exists("burden")) {
burden<-read_csv("data/seasonal_burden.csv") 
}

if(!exists("burden")) {
ed_omega<-read_csv("results/ed_omega.csv") 
}
```


### Predefine color palettes
```{r color-palettes}

col_diseases<-met.brewer("Hiroshige",2)

col_models<-met.brewer("Cassatt1",2)

col_season<-met.brewer("Hokusai1",6)
names(col_season)<-c("2017-2018", "2018-2019", "2019-2020", "2021-2022", "2022-2023", "2023-2024") 

col_subtype<-met.brewer("Isfahan2", 4)
                     
```

## Main Text Figures

### Fig 1. Intro to "Forecastability" 
#### v1.0
https://stackoverflow.com/questions/61761976/plotting-multiple-acf-with-ggplot
```{r fig1-v0, eval=FALSE}

flu_ts<-flu_hosp_full%>% filter(location=="us") %>% 
  arrange(time_value) %>%
  mutate(days= time_value-as.Date("2022-02-02")) 
flu_ts_state<-flu_hosp_full %>% filter(geo_value=="md") %>% 
  arrange(time_value) %>%
  mutate(days= time_value-as.Date("2022-02-02")) 

flu_ts %>% 
    pull(weekly_hosp) %>% ndiffs(test="kpss", alpha=confidence)
flu_ts %>% 
    pull(weekly_hosp) %>% ndiffs(test="adf", alpha=confidence)

flu_ts_state%>% 
    pull(weekly_hosp) %>% ndiffs(test="adf", alpha=confidence)
flu_ts_state%>% 
    pull(weekly_hosp) %>% ndiffs(test="kpss", alpha=confidence)

covid_ts<-covid_hosp_full %>% filter(geo_value=="us") %>% mutate(days= time_value-as.Date("2022-02-02")) %>% filter(days>=0)

flu_ts<- flu_ts %>% mutate(weekly_hosp_lag= weekly_hosp-lag(weekly_hosp)) %>% 
  mutate(weekly_hosp_lag2= weekly_hosp_lag-lag(weekly_hosp_lag)) %>% 
  drop_na()


flu_ts_state<-flu_ts_state %>% mutate(weekly_hosp_lag= weekly_hosp-lag(weekly_hosp)) %>%
  mutate(weekly_hosp_lag2= weekly_hosp_lag-lag(weekly_hosp_lag)) %>% 
  drop_na()

fore_flu<- as.numeric(Omega(flu_ts$weekly_hosp_lag, spectrum.control = sc, entropy.control = ec))
fore_flu_state<- as.numeric(Omega(flu_ts_state$weekly_hosp_lag, spectrum.control = sc, entropy.control = ec))

fore_covid<- combo %>% filter(season=="2022-2024", disease=="COVID-19", location=="us") %>% pull(omega)

#Random ts
days<-seq(1,nrow(flu_ts), by=1)
y<-rnorm(length(days),mean=1,sd=10)
random_ts<- data.frame(days,y)
random_ts %>%     
  pull(y) %>% ndiffs(test="kpss", alpha=confidence)
random_ts %>%     
  pull(y) %>% ndiffs(test="adf", alpha=confidence)
random_ts<- random_ts %>% mutate(lag_y= y-lag(y)) %>% 
  mutate(lay_y=lag_y-lag(lag_y)) %>% drop_na()


fore_rand<-as.numeric(Omega(random_ts$lag_y, spectrum.control = sc, entropy.control = ec))


Fig1A<- random_ts %>% ggplot(aes(days,y)) + geom_line() + theme_classic()+ ggtitle("Signal") +xlab("")+ylab("")
rand_acf<- data.frame(lag=acf(random_ts$y, main="", lag.max=365)$lag, acf=acf(random_ts$y, main="",lag.max=365)$acf) %>% filter(lag%%14 == 0)
Fig1B<- rand_acf %>% ggplot(aes(x=lag, y=acf)) +
  geom_segment(mapping = aes(xend = lag, yend = 0)) +
          geom_point()+ theme_classic() + ggtitle("Autocorrelation") +xlab("")+ylab("")
rand_spec<-data.frame(freq=mvspec(random_ts$y, log="no")$freq, spec= mvspec(random_ts$y, log="no")$spec) %>% 
  filter(freq<=0.025)

Fig1C<- rand_spec %>% ggplot(aes(x=freq, y=spec)) + geom_line() + theme_classic()+
  annotate(geom="text", x=0.02, y=250, label=paste(round(fore_rand,2),"%"),
              color="red")+
  ggtitle("Spectral Density") +xlab("")+ylab("")

# cowplot::plot_grid(A, B, C, ncol=1, label="AUTO")


#Flu state ts
Fig1D<- flu_ts_state %>% ggplot(aes(days, weekly_admits_p100K)) + geom_line()+ theme_classic() +xlab("")+ylab("")
#acf(flu_ts_state$weekly_admits_p100K, main="")
covid_acf<- data.frame(lag=acf(flu_ts_state$weekly_admits_p100K, lag.max=365)$lag, acf=acf(flu_ts_state$weekly_admits_p100K, lag.max=365)$acf) %>% filter(lag%%14 == 0)
Fig1E<- covid_acf %>% ggplot(aes(x=lag, y=acf)) +
  geom_segment(mapping = aes(xend = lag, yend = 0)) +
          geom_point()+ theme_classic() +xlab("")+ylab("")
flu_state_spec<- data.frame(freq=mvspec(flu_ts_state$weekly_admits_p100K, log="no")$freq, spec= mvspec(flu_ts_state$weekly_admits_p100K, log="no")$spec) %>% 
  filter(freq<=0.025)
Fig1F<- flu_state_spec %>% ggplot(aes(x=freq, y=spec)) + geom_line() + theme_classic()+
  annotate(geom="text", x=0.02, y=200, label=paste(round(fore_flu_state,2),"%"),
              color="red") +
  xlab("")+ylab("")


#Flu ts  
Fig1G<- flu_ts %>% ggplot(aes(days, weekly_admits_p100K)) + geom_line()+ theme_classic() +xlab("")+ylab("")
#acf(flu_ts$weekly_admits_p100K, main="")
flu_acf<- data.frame(lag=acf(flu_ts$weekly_admits_p100K, lag.max=365)$lag, acf=acf(flu_ts$weekly_admits_p100K, lag.max=365)$acf) %>% filter(lag%%14 == 0)
Fig1H<- flu_acf %>% ggplot(aes(x=lag, y=acf)) +
  geom_segment(mapping = aes(xend = lag, yend = 0)) +
          geom_point()+ theme_classic() +xlab("")+ylab("")
flu_spec<- data.frame(freq=mvspec(flu_ts$weekly_admits_p100K, log="no")$freq, spec= mvspec(flu_ts$weekly_admits_p100K, log="no")$spec) %>% 
  filter(freq<=0.025)
Fig1I<- flu_spec %>% ggplot(aes(x=freq, y=spec)) + geom_line() + theme_classic()+
  annotate(geom="text", x=0.02, y=300, label=paste(round(fore_flu,2),"%"),
              color="red") + 
  xlab("")+ylab("")

#Sine ts
sine_ts<- data.frame(days, y= sin(0.05*days))
sine_ts %>%     
  pull(y) %>% ndiffs(test="kpss", alpha=confidence)
sine_ts %>%     
  pull(y) %>% ndiffs(test="adf", alpha=confidence)
sine_ts<- sine_ts %>% mutate(lag_y= y-lag(y)) %>%
   # mutate(lag_y= lag_y-lag(lag_y)) %>% 
  drop_na()

fore_sine<- as.numeric(Omega(sine_ts$lag_y, spectrum.control = sc, entropy.control = ec))

Fig1J<- sine_ts %>% ggplot(aes(days,y)) + geom_line() + theme_classic() +xlab("Days")+ylab("")
acf(sine_ts$y, main="")
mvspec(sine_ts$y, log="no")
sine_acf<- data.frame(lag=acf(sine_ts$y, lag.max=365)$lag, acf=acf(sine_ts$y, lag.max=365)$acf) %>% filter(lag%%14 == 0)

Fig1K<- sine_acf %>% ggplot(aes(x=lag, y=acf)) +
  geom_segment(mapping = aes(xend = lag, yend = 0)) +
          geom_point()+ theme_classic() + xlab("Lag (Days)") + ylab("")
sine_spec<- data.frame(freq=mvspec(sine_ts$y, log="no")$freq, spec= mvspec(sine_ts$y, log="no")$spec) %>% 
  filter(freq<=0.025)
Fig1L<- sine_spec %>% ggplot(aes(x=freq, y=spec)) + geom_line() + theme_classic()+
  annotate(geom="text", x=0.02, y=60, label=paste(round(fore_sine,2),"%"),
              color="red")+
  xlab("Frequency")+ylab("")

cowplot::plot_grid(Fig1A, Fig1B, Fig1C, Fig1D, Fig1E, Fig1F, Fig1G, Fig1H, Fig1I, Fig1J, Fig1K, Fig1L, ncol=3,labels= "AUTO") # labels=c("A Signal", "B Autocorrelation", "C Spectral Density", "D", "E", "F", "G", "H", "I", "J", "K", "L"))


jpeg(file = "~/Projects/forecastability/Figs/Fig1.jpeg",  
    width = 10, # The width of the plot in inches
    height = 8, units="in", res=500) # The height of the plot in inches

cowplot::plot_grid(Fig1A, Fig1B, Fig1C, Fig1D, Fig1E, Fig1F, Fig1G, Fig1H, Fig1I, Fig1J, Fig1K, Fig1L, ncol=3,labels= "AUTO") # labels=c("A Signal", "B Autocorrelation", "C Spectral Density", "D", "E", "F", "G", "H", "I", "J", "K", "L"))

dev.off()

```
#### v2.0
```{r fig1-v1}
#Random ts
days<-seq(1,365, by=1)
y<-rnorm(length(days),mean=0,sd=1)
random_ts<- data.frame(days,y)

random_ts %>%     
  pull(y) %>% ndiffs(test="kpss", alpha=confidence)
random_ts %>%     
  pull(y) %>% ndiffs(test="adf", alpha=confidence)
random_ts<- random_ts %>% 
  mutate(lag_y= y-lag(y)) %>% 
  drop_na()

fore_rand<-as.numeric(Omega(random_ts$lag_y, spectrum.control = sc, entropy.control = ec))


Fig1A<- random_ts %>% ggplot(aes(days,y)) + geom_line() + theme_classic()+ ggtitle("Signal") +xlab("")+ylab("")
rand_acf<- data.frame(lag=acf(random_ts$y, main="", lag.max=365)$lag, acf=acf(random_ts$y, main="",lag.max=365)$acf) %>% filter(lag%%14 == 0)

Fig1B<- rand_acf %>% ggplot(aes(x=lag, y=acf)) +
  geom_segment(mapping = aes(xend = lag, yend = 0)) +
          geom_point()+ theme_classic() + ggtitle("Autocorrelation") +xlab("")+ylab("")
rand_spec<-data.frame(freq=mvspec(random_ts$y, log="no")$freq, spec= mvspec(random_ts$y, log="no")$spec) %>% 
  filter(freq<=0.05)

Fig1C<- rand_spec %>% ggplot(aes(x=freq, y=spec)) + geom_line() + theme_classic()+
  annotate(geom="text", x=0.02, y=2.5, label=paste(round(fore_rand,2),"%"),
              color="red")+
  ggtitle("Spectral Density") +xlab("")+ylab("")

#SIR simulation
# Parameters and initial conditions.
parameters <- c(beta = 0.2, gamma = 0.1)
initials <- c(S = 1 - 1e-06, I = 1e-06, R = 1 - (1 - 1e-06 - 1e-06))
# Solve and plot.
sir <- SIR(pars = parameters, init = initials, time = 0:365)
PlotMods(sir)

sir$results$I %>% ndiffs(test="kpss", alpha=confidence)
sir$results$I %>% ndiffs(test="adf", alpha=confidence)

sir_results<- data.frame(days=sir$results$time, y=sir$result$I) %>% mutate(lag_y= sir$results$I- lag(sir$results$I)) 

# fore_sir <- Omega(sir$results$I, spectrum.control = sc, entropy.control = ec)
fore_sir <- Omega(na.omit(sir_results$lag_y), spectrum.control = sc, entropy.control = ec)

Fig1D<- sir_results %>% ggplot(aes(days,y)) + geom_line() + theme_classic()+ ggtitle("") +xlab("")+ylab("")

sir_acf<- data.frame(lag=acf(sir_results$y, main="", lag.max=365)$lag, acf=acf(sir_results$y, main="",lag.max=365)$acf) %>% filter(lag%%14 == 0)

Fig1E<- sir_acf %>% ggplot(aes(x=lag, y=acf)) +
  geom_segment(mapping = aes(xend = lag, yend = 0)) +
          geom_point()+ theme_classic() + ggtitle("Autocorrelation") +xlab("")+ylab("")

sir_spec<-data.frame(freq=mvspec(sir_results$y, log="no")$freq, spec= mvspec(sir_results$y, log="no")$spec) %>% 
  filter(freq<=0.05)

Fig1F<- sir_spec %>% ggplot(aes(x=freq, y=spec)) + geom_line() + theme_classic()+
  annotate(geom="text", x=0.02, y=0.1, label=paste(round(fore_sir,2),"%"),
              color="red")+
  ggtitle("Spectral Density") +xlab("")+ylab("")


#Sine ts
sine_ts<- data.frame(days, y= sin(0.05*days))
sine_ts %>%     
  pull(y) %>% ndiffs(test="kpss", alpha=confidence)
sine_ts %>%     
  pull(y) %>% ndiffs(test="adf", alpha=confidence)
sine_ts<- sine_ts %>% mutate(lag_y= y-lag(y)) %>%
   # mutate(lag_y= lag_y-lag(lag_y)) %>% 
  drop_na()

fore_sine<- as.numeric(Omega(sine_ts$lag_y, spectrum.control = sc, entropy.control = ec))

Fig1G<- sine_ts %>% ggplot(aes(days,y)) + geom_line() + theme_classic() +xlab("Days")+ylab("")
acf(sine_ts$y, main="")
mvspec(sine_ts$y, log="no")
sine_acf<- data.frame(lag=acf(sine_ts$y, lag.max=365)$lag, acf=acf(sine_ts$y, lag.max=365)$acf) %>% filter(lag%%14 == 0)

Fig1H<- sine_acf %>% ggplot(aes(x=lag, y=acf)) +
  geom_segment(mapping = aes(xend = lag, yend = 0)) +
          geom_point()+ theme_classic() + xlab("Lag (Days)") + ylab("")
sine_spec<- data.frame(freq=mvspec(sine_ts$y, log="no")$freq, spec= mvspec(sine_ts$y, log="no")$spec) %>% 
  filter(freq<=0.05)
Fig1I<- sine_spec %>% ggplot(aes(x=freq, y=spec)) + geom_line() + theme_classic()+
  annotate(geom="text", x=0.02, y=60, label=paste(round(fore_sine,2),"%"),
              color="red")+
  xlab("Frequency")+ylab("")

cowplot::plot_grid(Fig1A, Fig1B, Fig1C, Fig1D, Fig1E, Fig1F, Fig1G, Fig1H, Fig1I, ncol=3,labels= "AUTO") # labels=c("A Signal", "B Autocorrelation", "C Spectral Density", "D", "E", "F", "G", "H", "I", "J", "K", "L"))


jpeg(file = "~/Projects/forecastability/Figs/Fig1.jpeg",  
    width = 10, # The width of the plot in inches
    height = 8, units="in", res=500) # The height of the plot in inches

cowplot::plot_grid(Fig1A, Fig1B, Fig1C, Fig1D, Fig1E, Fig1F, Fig1G, Fig1H, Fig1I, Fig1J, Fig1K, Fig1L, ncol=3,labels= "AUTO") # labels=c("A Signal", "B Autocorrelation", "C Spectral Density", "D", "E", "F", "G", "H", "I", "J", "K", "L"))

dev.off()
```




### Fig 2. HCAI

```{r fig2}
Fig2A<- hcai_combined %>% filter(season %in% c("2017-2018", "2018-2019", "2019-2020", "2021-2022")) %>% ggplot(aes(x=log(pop), y=omega, col=season)) +
  geom_point()+ 
  theme_classic()+  
  # geom_text_repel(aes(label=location))+ 
  # ggtitle("HCAI Weekly Influenza Admissions Time Series Forecastability (County and State)")+
  xlab("Natural logarithm of population size")+
  ylab(expression(paste("Forecastability (",  Omega, ") (%)")))+
  geom_smooth(aes(fill=season), method="gam", formula= y~s(x, bs = "cs"))+
  scale_color_manual(name = "Season", values=col_season)+
  scale_fill_manual(name = "Season", values=col_season)+
  theme(legend.position="bottom")+
  # xlim(c(9,20))+
  ylim(c(0,45))


jpeg(file = "~/Projects/forecastability/Figs/Fig2a.jpeg",  
    width = 10, # The width of the plot in inches
    height = 8, units="in", res=500) # The height of the plot in inches

Fig2A

dev.off()


Fig2B<- hcai_state_omega %>% mutate(season_dummy=season) %>% 
  separate(season_dummy, into = c("season_start", "season_end"), sep = "-") %>% 
  left_join(hcai_burden, by=c("location", "season")) %>% 
  left_join(flu_subtypes, by="season")%>%
  filter(location=="California", season %notin% c("1999-2000", "2020-2021", "2022-2023")) %>% ggplot(aes(x=log(max_admits), y=omega)) + 
  geom_smooth(method = "lm", se = TRUE, alpha=0.3)+
  geom_point(aes(col=CDPH), size=2)+ 
  geom_text_repel(aes(label=season_start, col=CDPH))+ 
  theme_classic()+
  xlab("Natural logarithm of peak weekly admissions")+
  ylab(expression(paste("Forecastability (",  Omega, ") (%)")))+
  scale_color_manual(name = "Predominant Influenza Subtype", values=col_subtype)+
  theme(legend.position="bottom")+
  ylim(c(0,45))

Fig2_inset<- hcai_combined %>% filter(season %in% c("2017-2018", "2018-2019", "2019-2020", "2021-2022")) %>% ggplot(aes(x=log(pop), y=omega, col=season)) +
  geom_point()+ 
  theme_classic()+  
  geom_smooth(aes(fill=season), method="gam", formula= y~s(x, bs = "cs"))+
  scale_color_manual(name = "Season", values=col_season)+
  scale_fill_manual(name = "Season", values=col_season)+
  theme(legend.position="none", axis.title = element_blank())+
  xlim(c(10,14))+
  ylim(c(8,22))


Fig2_main<- cowplot::plot_grid(Fig2A, Fig2B, ncol=1, labels="AUTO", align="h")
Fig2 <- ggdraw(Fig2_main) +
      draw_plot(Fig2_inset, x = 0.1, y = 0.75, width = 0.35, height = 0.25)
```

### Fig 3. Population size
```{r}
# Forecastability scores
omega_flu_us<- forecastability %>% filter(location=="us", season=="2022-2023", disease=="Influenza") %>% pull(omega) %>% unique(.)

omega_covid_us<- forecastability %>% filter(location=="us", season=="2022-2023", disease=="COVID-19") %>% pull(omega) %>% unique(.)

omega_flu_wy<- forecastability %>% filter(location=="wy", season=="2022-2023", disease=="Influenza") %>% pull(omega) %>% unique(.)

omega_covid_wy<- forecastability %>% filter(location=="wy", season=="2022-2023", disease=="COVID-19") %>% pull(omega) %>% unique(.)

omega_labels<- forecastability %>% filter(location %in% c("wy", "us"), season=="2022-2023", disease %in% c("COVID-19", "Influenza")) %>% dplyr::select(location, season, disease, omega) %>% unique(.) %>%
  mutate(geography= c("Large population size", "Large population size", "Small population size", "Small population size"))

## Influenza: State time series- small scale
flu_ts_state<-flu_hosp_full %>% filter(geo_value=="wy", time_value >= season22_start_MMWR & time_value < season23_start_MMWR) %>% 
  arrange(time_value) 

# Fig3A <- flu_ts_state %>% ggplot(aes(time_value, weekly_admits_p100K)) + geom_line()+ theme_classic() +xlab("")+ylab("")
# #acf(flu_ts_state$weekly_admits_p100K, main="")

flu_state_spec<- data.frame(freq=mvspec(flu_ts_state$weekly_admits_p100K, log="no")$freq, spec= mvspec(flu_ts_state$weekly_admits_p100K, log="y")$spec) %>% 
  filter(freq<=0.1)

# Fig3B <- flu_state_spec %>% ggplot(aes(x=freq, y=spec)) + geom_line() + theme_classic()+
#   annotate(geom="text", x=0.02, y=200, label=paste(round(omega_flu_wy,2),"%"),
#               color="red") +
#   xlab("")+ylab("")+
#   # scale_y_log10()+
#   geom_hline(yintercept=1, lty="dashed")

## Influenza: US time series- large scale
flu_ts<-flu_hosp_full%>% filter(geo_value=="us", time_value >= season22_start_MMWR & time_value < season23_start_MMWR) %>% 
  arrange(time_value) 

# Fig3C<- flu_ts %>% ggplot(aes(time_value, weekly_admits_p100K)) + geom_line()+ theme_classic() +xlab("")+ylab("")

flu_us_spec<- data.frame(freq=mvspec(flu_ts$weekly_admits_p100K)$freq, spec= mvspec(flu_ts$weekly_admits_p100K)$spec) %>% 
  filter(freq<=0.1)

# Fig3D<- flu_us_spec %>% ggplot(aes(x=freq, y=spec)) + geom_line() + theme_classic()+
#   annotate(geom="text", x=0.02, y=200, label=paste(round(omega_flu_us,2),"%"),
#               color="red") +
#   xlab("")+ylab("")+
#   # scale_y_log10()+
#   geom_hline(yintercept=1, lty="dashed")


#COVID-19: state/small scale

covid_ts_state<-covid_hosp_full %>% filter(geo_value=="wy",  time_value >= season22_start_MMWR & time_value < season23_start_MMWR) %>% 
  arrange(time_value) 
# 
# Fig3E<- covid_ts_state %>% ggplot(aes(time_value, weekly_admits_p100K)) + geom_line()+ theme_classic() +xlab("")+ylab("")
# #acf(flu_ts_state$weekly_admits_p100K, main="")

covid_state_spec<- data.frame(freq=mvspec(covid_ts_state$weekly_admits_p100K, log="no")$freq, spec= mvspec(covid_ts_state$weekly_admits_p100K, log="no")$spec) %>% 
  filter(freq<=0.1)

# Fig3F<- covid_state_spec %>% ggplot(aes(x=freq, y=spec)) + geom_line() + theme_classic()+
#   annotate(geom="text", x=0.02, y=200, label=paste(round(omega_covid_wy,2),"%"),
#               color="red") +
#   xlab("")+ylab("")+
#   # scale_y_log10()+
#   geom_hline(yintercept=1, lty="dashed")

#COVID-19: US/large scale

covid_ts<-covid_hosp_full%>% filter(geo_value=="us",  time_value >= season22_start_MMWR & time_value < season23_start_MMWR) %>% 
  arrange(time_value)

# Fig3G<- covid_ts %>% ggplot(aes(time_value, weekly_admits_p100K)) + geom_line()+ theme_classic() +xlab("")+ylab("")

covid_us_spec<- data.frame(freq=mvspec(covid_ts$weekly_admits_p100K, log="no")$freq, spec= mvspec(covid_ts$weekly_admits_p100K, log="yes")$spec) %>%
  filter(freq<=0.1)

# Fig3H<- covid_us_spec %>% ggplot(aes(x=freq, y=spec)) + geom_line() + theme_classic()+
#   annotate(geom="text", x=0.02, y=200, label=paste(round(omega_covid_us,2),"%"),
#               color="red") +
#   xlab("")+ylab("")+
#   # scale_y_log10()+
#   geom_hline(yintercept=1, lty="dashed")

# cowplot::plot_grid(Fig3A, Fig3B, Fig3C, Fig3D, Fig3E, Fig3F, Fig3G, Fig3H, ncol=2, labels="AUTO")

Fig3A<-combo %>% filter(season %in% c("2022-2023", "2023-2024"), location_level=="state") %>%
  ggplot(aes(x=log_pop, y=omega, col=season))  +
  geom_point(alpha=0.5) +
  facet_grid(.~disease) +
  # geom_line(data = flu_piece, color = col_diseases[2])+
  # geom_line(data = covid_piece, color = col_diseases[1])+
  geom_smooth(aes(fill=season, col=season), method="gam", formula= y~s(x, bs = "cs"))+

  xlab("Natural logarithm of population size")+
  ylab(expression(paste("Forecastability (",  Omega, ") (%)")))+
    scale_color_manual(name = "Season", values=col_season)+
      scale_fill_manual(name = "Season", values=col_season)+
  theme_classic()

spec_all<-covid_us_spec %>% mutate(location="US", disease="COVID-19", geography= "Large population size", omega=omega_covid_us) %>%
  bind_rows(covid_state_spec %>% mutate(location="WY", disease= "COVID-19", geography= "Small population size",omega=omega_covid_wy)) %>% 
  bind_rows(flu_us_spec %>% mutate(location="US", disease= "Influenza", geography= "Large population size", omega=omega_flu_us)) %>%
  bind_rows(flu_state_spec %>% mutate(location="WY", disease= "Influenza", geography= "Small population size", omega=omega_flu_wy))

Fig3C<- spec_all %>% ggplot(aes(x=freq, y=spec)) + 
  facet_grid(disease ~ geography) + 
  geom_line() + 
  theme_classic()+
  theme(axis.text.x = element_text(angle=45, hjust = 1, vjust = 1))+
  xlab("Frequency")+
  ylab("Spectral Density")+
  geom_text(data=omega_labels,
    aes(x = 0.075, y=350, label = paste(round(omega,2),"%")),
    hjust = 0,    # Horizontal justification (0 = left, 1 = right)
    vjust = 1,    # Vertical justification (1 = top, 0 = bottom)
    size = 4,     # Font size
    color = "red"
  )

ts_all<-covid_ts %>% mutate(location="US", disease="COVID-19", geography= "Large population size") %>%
  bind_rows(covid_ts_state %>% mutate(location="WY", disease= "COVID-19", geography= "Small population size")) %>% bind_rows(flu_ts %>% mutate(location="US", disease= "Influenza", geography= "Large population size")) %>%
  bind_rows(flu_ts_state %>% mutate(location="WY", disease= "Influenza", geography= "Small population size"))

Fig3B<- ts_all %>% ggplot(aes(time_value, weekly_admits_p100K)) + 
  geom_line()+ 
  theme_classic()+  
  facet_grid(disease ~ geography)+
  xlab("Date")+
  ylab("Weekly Admissions/100K")+
  theme(axis.text.x = element_text(angle=45, hjust = 1, vjust = 1))

cowplot::plot_grid(Fig3A, Fig3B, Fig3C, ncol=1, labels="AUTO")
```

### Fig 4. Forecast Performance
```{r fig4}

Fig4A<-forecastability %>% 
  dplyr::filter(model %in% c("Baseline", "Ensemble")) %>%
  ggplot(aes(x=omega, y=ae_median, col=model)) + 
  facet_grid(season~disease) + 
  geom_point() +
   geom_smooth(aes(fill=model), method = "lm", se = TRUE)+
   theme_classic()+
# theme(legend.position = "none")+
    xlab(expression(paste("Forecastability (",  Omega, ") (%)")))+
  ylab("MAE on Log Scale")+
    scale_color_manual(name = "Model", values=col_models)+
      scale_fill_manual(name = "Model", values=col_models)


Fig4B<-forecastability %>%
  dplyr::filter(model %in% c("Baseline", "Ensemble")) %>%  ggplot(aes(x=omega, y=wis, col=model)) + facet_grid(season~disease) + 
  geom_point() +
   geom_smooth(aes(fill=model), method = "lm", se = TRUE)+
   theme_classic()+
# theme(legend.position = "none")+
      xlab(expression(paste("Forecastability (",  Omega, ") (%)")))+
  ylab("Mean WIS on Log Scale")+
    scale_color_manual(name = "Model", values=col_models)+
      scale_fill_manual(name = "Model", values=col_models)

Fig4C<- forecastability %>%
  dplyr::filter(model == "Ensemble") %>%  ggplot(aes(x=omega, y=wis_scaled_relative_skill, col=season)) + 
  facet_grid(.~ disease) + 
  geom_point() +
   geom_smooth(aes(fill=season), method = "lm", se = TRUE)+
   theme_classic()+
    scale_color_manual(name = "Season", values=col_season)+
      scale_fill_manual(name = "Season", values=col_season)+
# theme(legend.position = "none")+
      xlab(expression(paste("Forecastability (",  Omega, ") (%)")))+
  ylab("Scaled Relative Skill")


Fig4 <-cowplot::plot_grid(Fig4A, Fig4B, Fig4C, ncol=1, labels="AUTO")

Fig4
```

### Fig 5. Horizon
```{r Fig5-horizon}

Fig5<-forecastability_horizon %>%
  dplyr::filter(model == "Ensemble") %>%  ggplot(aes(x=omega, y=wis_scaled_relative_skill, col=season)) + 
  facet_grid(horizon~ disease) + 
  geom_point() +
   geom_smooth(aes(fill=season), method = "lm", se = TRUE)+
   theme_classic()+
    scale_color_manual(name = "Season", values=col_season)+
      scale_fill_manual(name = "Season", values=col_season)+
# theme(legend.position = "none")+
      xlab(expression(paste("Forecastability (",  Omega, ") (%)")))+
  ylab("Scaled Relative Skill")

Fig5

```

### Fig 6. Sampling Frequency
```{r fig6}
cadence<- state_flu_cadence %>% bind_rows(state_covid_cadence)

Fig6<- cadence %>% filter(season %notin% c("2022-2024")) %>% 
  ggplot(aes(x=log(pop), y=omega, col=season)) +
  geom_point()+ 
  facet_grid( disease ~cadence)+
  theme_classic()+  
  geom_smooth(aes(fill=season), method = "lm", se = TRUE, alpha=0.3)+
  geom_text_repel(aes(label=state))+ 
  scale_color_manual(name = "Season", values=col_season)+
  scale_fill_manual(name = "Season", values=col_season)+
  ylim(c(0,45))+
  theme(legend.position="bottom")+
  xlab("Natural logarithm of population size")+
  ylab(expression(paste("Forecastability (",  Omega, ") (%)")))

Fig6


# SupFig2<-state_covid_cadence %>% 
#   filter(season != "2022-2024")%>%
#   ggplot(aes(x=log(pop), y=omega, col=season)) +
#   facet_wrap(data_type ~.)+
#   geom_point()+ 
#   theme_classic()+  
#   geom_smooth(aes(fill=season), method = "lm", se = TRUE, alpha=0.3)+
#   geom_text_repel(aes(label=state))+ 
#   scale_color_manual(name = "Season", values=col_season)+
#   scale_fill_manual(name = "Season", values=col_season)+
#   ylim(c(0,45))+
#   theme(legend.position="bottom")+
#   xlab("Natural logarithm of population size")+
#   ylab(expression(paste("Forecastability (",  Omega, ") (%)")))


```


### Fig 7. Temporal Aggregation

Effects of data aggregation/smoothing (rolling sum)

```{r fig7}

### Single state and different intervals of rolling sum ----------
library(covidcast)
hosp_flu <-covidcast_signal(data_source = "hhs",
                          signal = "confirmed_admissions_influenza_1d",
                          start_day = "2022-02-02", end_day = Sys.Date(), geo_type=c("state"), geo_value=c("md"))
min_date<- min(hosp_flu$time_value)
md_flu<- hosp_flu %>%   arrange(time_value) %>% mutate(days=time_value-min_date)

# fore_md_rolling<-data.frame(sampling_frequency=integer(365), forecastability=numeric(365))
# 
# for (i in 1:56){
#   md_temp<-md_flu %>% mutate(days=as.numeric(days)) %>% mutate(rolling_sum= zoo::rollsum(value, k = i, align = "right", fill = NA)) %>% 
#     filter(days %% i==0)  %>% 
#     mutate(lag_y= rolling_sum-lag(rolling_sum)) %>% drop_na(lag_y)
#   fore_flu_temp<- as.numeric(Omega(md_temp$lag_y, spectrum.control = sc, entropy.control = ec))
#   fore_md_rolling$sampling_frequency[i]=i
#   fore_md_rolling$forecastability[i]=fore_flu_temp
# }
# 
# md_temp %>% ggplot(aes(days,rolling_sum)) + geom_point()+geom_line() + theme_classic() +xlab("Days")+ylab("")
# 
# 
# Fig5D<- fore_md_rolling %>% filter(sampling_frequency !=0) %>% ggplot(aes(sampling_frequency, forecastability))+ geom_point() + theme_classic()+
#     geom_smooth(aes(sampling_frequency, forecastability), method="gam", formula= y~s(x, bs = "cs"))+ 
#    xlab("Days of Aggregation")+
#   ylab("Forecastabilty")
# 
# Fig5A<-md_flu %>% mutate(days=as.numeric(days)) %>% mutate(rolling_sum= value)  %>% mutate(lag_y= rolling_sum-lag(rolling_sum)) %>% drop_na(lag_y) %>% ggplot(aes(days,rolling_sum)) +geom_line() + theme_classic() +xlab("Days")+ylab("Incident Hospital Admissions")+
#   ggtitle("Daily Reporting") 
# 
# Fig5B<-md_flu %>% mutate(days=as.numeric(days)) %>% mutate(rolling_sum= zoo::rollsum(value, k = 28, align = "right", fill = NA)) %>% filter(days %% 28==0)  %>% mutate(lag_y= rolling_sum-lag(rolling_sum)) %>% drop_na(lag_y) %>% ggplot(aes(days,rolling_sum)) +geom_line() + geom_point()+ theme_classic() +xlab("Days")+ylab("")+
#   ggtitle("4 Week Aggregation")
# 
# Fig5C<- md_flu %>% mutate(days=as.numeric(days)) %>% mutate(rolling_sum= zoo::rollsum(value, k = 56, align = "right", fill = NA)) %>% filter(days %% 56==0)  %>% mutate(lag_y= rolling_sum-lag(rolling_sum)) %>% drop_na(lag_y) %>% ggplot(aes(days,rolling_sum)) +geom_line() + geom_point()+ theme_classic() +xlab("Days")+ylab("")+
#   ggtitle("8 Week Aggregation")
# 
# Fig5_top<-cowplot::plot_grid(Fig5A, Fig5B, Fig5C, nrow=1, labels="AUTO")
# Fig5<-cowplot::plot_grid(Fig5_top, Fig5D, nrow=2, labels=c("","D"))
# Fig5

### alternatively-- rolling/smoothing without downsampling 
fore_md_rolling<-data.frame(sampling_frequency=integer(365), forecastability=numeric(365))

for (i in 1:56){
  md_temp<-md_flu %>% mutate(days=as.numeric(days)) %>% mutate(rolling_sum= zoo::rollsum(value, k = i, align = "right", fill = NA)) %>% 
    mutate(lag_y= rolling_sum-lag(rolling_sum)) %>% drop_na(lag_y)
  fore_flu_temp<- as.numeric(Omega(md_temp$lag_y, spectrum.control = sc, entropy.control = ec))
  fore_md_rolling$sampling_frequency[i]=i
  fore_md_rolling$forecastability[i]=fore_flu_temp
}

md_temp %>% ggplot(aes(days,rolling_sum)) + geom_point()+geom_line() + theme_classic() +xlab("Days")+ylab("")


Fig5D<- fore_md_rolling %>% filter(sampling_frequency !=0) %>% ggplot(aes(sampling_frequency, forecastability))+ geom_point() + theme_classic()+
    geom_smooth(aes(sampling_frequency, forecastability), method="gam", formula= y~s(x, bs = "cs"))+ 
   xlab("Rolling Sum (Days)")+
   ylab(expression(paste("Forecastability (",  Omega, ") (%)")))


Fig7A<-md_flu %>% mutate(days=as.numeric(days)) %>% mutate(rolling_sum= value)  %>% mutate(lag_y= rolling_sum-lag(rolling_sum)) %>% drop_na(lag_y) %>% ggplot(aes(days,rolling_sum)) +geom_line() + theme_classic() +xlab("Days")+ylab("Incident Hospital Admissions")+
  ggtitle("Daily Reporting") 

Fig7B<-md_flu %>% mutate(days=as.numeric(days)) %>% mutate(rolling_sum= zoo::rollsum(value, k = 28, align = "right", fill = NA))   %>% mutate(lag_y= rolling_sum-lag(rolling_sum)) %>% drop_na(lag_y) %>% ggplot(aes(days,rolling_sum)) +geom_line() +  theme_classic() +xlab("Days")+ylab("")+
  ggtitle("28 Day Rolling Sum")

Fig7C<- md_flu %>% mutate(days=as.numeric(days)) %>% mutate(rolling_sum= zoo::rollsum(value, k = 56, align = "right", fill = NA))  %>% mutate(lag_y= rolling_sum-lag(rolling_sum)) %>% drop_na(lag_y) %>% ggplot(aes(days,rolling_sum)) +geom_line() +  theme_classic() +xlab("Days")+ylab("")+
  ggtitle("56 Day Rolling Sum")

Fig7_top<-cowplot::plot_grid(Fig7A, Fig7B, Fig7C, nrow=1, labels="AUTO")
Fig7<-cowplot::plot_grid(Fig7_top, Fig7D, nrow=2, labels=c("","D"))
Fig75


```






## Supplementary Figs

### Fig S1. HCAI

```{r SupFig1}
SupFig1<-hcai_state_omega %>% mutate(season_dummy=season) %>% 
  separate(season_dummy, into = c("season_start", "season_end"), sep = "-") %>%  
  left_join(hcai_burden, by=c("location", "season")) %>% 
    left_join(flu_subtypes, by="season")%>%
  filter(location=="California", season %notin% c("1999-2000", "2020-2021", "2022-2023")) %>% ggplot(aes(x=log(total_admits), y=omega)) + 
  geom_smooth(method = "lm", se = TRUE, alpha=0.3)+
  geom_text_repel(aes(label=season_start, col=CDPH))+ 
  geom_point(aes(col=CDPH), size=2)+ 
  theme_classic()+
  xlab("Natural logarithm of cumulative hospital admissions")+
  ylab(expression(paste("Forecastability (",  Omega, ") (%)")))+
  scale_color_manual(name = "Predominant Influenza Subtype", values=col_subtype)

SupFig1


```

### Fig S2. %ED Visits
ED visits vs. forecastability
```{r ed-omega}
ed_omega %>% ggplot(aes(log(pop), omega)) + 
  geom_point() + 
  facet_grid(season~disease)+ 
  geom_smooth(method="lm")+
  theme_classic()+
  xlab("Natural log of population size")+
  ylab(expression(paste("Forecastability (",  Omega, ") (%)")))
```

### Fig S3. Pop size
```{r}
SupFig2A<-forecastability %>% 
  dplyr::filter(model %in% c("Baseline", "Ensemble")) %>%
  ggplot(aes(x=log(pop), y=ae_median, col=model)) + 
  facet_grid(season~disease) + 
  geom_point() +
   geom_smooth(aes(fill=model), method = "lm", se = TRUE)+
   theme_classic()+
# theme(legend.position = "none")+
    xlab("Natural log of population size")+
  ylab("MAE on Log Scale")+
    scale_color_manual(name = "Model", values=col_models)+
      scale_fill_manual(name = "Model", values=col_models)


SupFig2B<-forecastability %>%
  dplyr::filter(model %in% c("Baseline", "Ensemble")) %>%  ggplot(aes(x=log(pop), y=wis, col=model)) + facet_grid(season~disease) + 
  geom_point() +
   geom_smooth(aes(fill=model), method = "lm", se = TRUE)+
   theme_classic()+
# theme(legend.position = "none")+
    xlab("Natural log of population size")+
  ylab("Mean WIS on Log Scale")+
    scale_color_manual(name = "Model", values=col_models)+
      scale_fill_manual(name = "Model", values=col_models)

SupFig2C<- forecastability %>%
  dplyr::filter(model == "Ensemble") %>%  ggplot(aes(x=log(pop), y=wis_scaled_relative_skill, col=season)) + 
  facet_grid(.~ disease) + 
  geom_point() +
   geom_smooth(aes(fill=season), method = "lm", se = TRUE)+
   theme_classic()+
    scale_color_manual(name = "Season", values=col_season)+
      scale_fill_manual(name = "Season", values=col_season)+
# theme(legend.position = "none")+
    xlab("Natural log of population size")+
  ylab("Scaled Relative Skill")


SupFig2 <-cowplot::plot_grid(SupFig2A, SupFig2B, SupFig2C, ncol=1, labels="AUTO")

SupFig2
```


### Fig S4. Individual models
```{r sup-fig4}


SupFig2A<-forecastability %>% 
  dplyr::filter(model %notin% c("MOBS-GLEAM", "CEPH-Rtrend")) %>%
  ggplot(aes(x=log(pop), y=ae_median, col=model))+ facet_grid(season~disease)+ geom_point()+
   theme_classic()+
   geom_smooth(method = "lm", se = TRUE)+
theme(legend.position = "none")+
  xlab("ln(Population size)")+
  ylab("Mean Absolute Error")


SupFig2B<-forecastability %>% 
  dplyr::filter(model %notin% c("MOBS-GLEAM", "CEPH-Rtrend")) %>%
  ggplot(aes(x=log(pop), y=wis, col=model))+ facet_grid(season~disease)+ geom_point()+
   theme_classic()+
   geom_smooth(method = "lm", se = TRUE)+
    xlab("ln(Population size)")+
theme(legend.position = "none")+
  ylab("Mean Weighted Interval Score")


legend_plot<- forecastability %>% 
    dplyr::filter(model %notin% c("MOBS-GLEAM", "CEPH-Rtrend")) %>%
  ggplot(aes(x=log(pop), y=ae_median, col=model))+ facet_wrap(season~disease)+ geom_point()+
   theme_classic()+
   geom_smooth(method = "lm", se = TRUE)+
    xlab("ln(Population size)")+
  ylab("Median Absolute Error")+
  theme(legend.position="bottom")

SupFig2_legend <- get_plot_component(legend_plot, "guide-box", return_all = TRUE)

preSupFig2<-cowplot::plot_grid(SupFig2A, SupFig2B, nrow=1, labels="AUTO")
SupFig2<-cowplot::plot_grid(preSupFig2, SupFig2_legend[[3]], nrow=2, rel_heights = c(2,0.5))


SupFig2
```

### Fig S5. Cumulative incidence
```{r}
SupFigA<- combo %>% left_join(burden, by=c("disease", "location", "season")) %>% 
  filter(season %in% c("2022-2023", "2023-2024"), location_level=="state") %>%
  ggplot(aes(x=log(burden), y=omega, col=season))  +
  geom_point(alpha=0.5) +
  facet_grid(.~disease) +
  # geom_line(data = flu_piece, color = col_diseases[2])+
  # geom_line(data = covid_piece, color = col_diseases[1])+
  geom_smooth(aes(fill=season, col=season), method="gam", formula= y~s(x, bs = "cs"))+

  xlab("Natural logarithm of cumulative incidence")+
  ylab(expression(paste("Forecastability (",  Omega, ") (%)")))+
    scale_color_manual(name = "Season", values=col_season)+
      scale_fill_manual(name = "Season", values=col_season)+
  theme_classic()

SupFigB<- forecastability %>% 
  dplyr::filter(model %in% c("Baseline", "Ensemble"), location %notin% c("as", "vi")) %>%
  ggplot(aes(x=log(burden), y=ae_median, col=model)) + 
  facet_grid(season~disease) + 
  geom_point() +
   geom_smooth(aes(fill=model), method = "lm", se = TRUE)+
   theme_classic()+
# theme(legend.position = "none")+
      xlab("Natural log of cumulative incidence")+
  ylab("MAE on Log Scale")+
    scale_color_manual(name = "Model", values=col_models)+
      scale_fill_manual(name = "Model", values=col_models)

SupFigC<- forecastability %>% 
  dplyr::filter(model %in% c("Baseline", "Ensemble"), location %notin% c("as", "vi")) %>%
  ggplot(aes(x=log(burden), y=wis_relative_skill, col=model)) + 
  facet_grid(season~disease) + 
  geom_point() +
   geom_smooth(aes(fill=model), method = "lm", se = TRUE)+
   theme_classic()+
# theme(legend.position = "none")+
      xlab("Natural log of cumulative incidence")+
  ylab("WIS on Log Scale")+
    scale_color_manual(name = "Model", values=col_models)+
      scale_fill_manual(name = "Model", values=col_models)

SupFigD<- forecastability %>%
  dplyr::filter(model == "Ensemble", location %notin% c("as", "vi")) %>%  ggplot(aes(x=log(burden), y=wis_scaled_relative_skill, col=season)) + 
  facet_grid(.~ disease) + 
  geom_point() +
   geom_smooth(aes(fill=season), method = "lm", se = TRUE)+
   theme_classic()+
    scale_color_manual(name = "Season", values=col_season)+
      scale_fill_manual(name = "Season", values=col_season)+
# theme(legend.position = "none")+
      xlab("Natural log of cumulative incidence")+
  ylab("Scaled Relative Skill")

cowplot::plot_grid(SupFigA, SupFigB, SupFigC, SupFigD, ncol=1)
```


### Fig S6. Forecast Performance (Summer season July/June)
```{r figS6}

FigS6A<-forecastability_summer %>% 
  dplyr::filter(model %in% c("Baseline", "Ensemble")) %>%
  ggplot(aes(x=omega, y=ae_median, col=model)) + 
  facet_grid(season~disease) + 
  geom_point() +
   geom_smooth(aes(fill=model), method = "lm", se = TRUE)+
   theme_classic()+
# theme(legend.position = "none")+
    xlab(expression(paste("Forecastability (",  Omega, ") (%)")))+
  ylab("MAE on Log Scale")+
    scale_color_manual(name = "Model", values=col_models)+
      scale_fill_manual(name = "Model", values=col_models)


FigS6B<-forecastability %>%
  dplyr::filter(model %in% c("Baseline", "Ensemble")) %>%  ggplot(aes(x=omega, y=wis, col=model)) + facet_grid(season~disease) + 
  geom_point() +
   geom_smooth(aes(fill=model), method = "lm", se = TRUE)+
   theme_classic()+
# theme(legend.position = "none")+
      xlab(expression(paste("Forecastability (",  Omega, ") (%)")))+
  ylab("Mean WIS on Log Scale")+
    scale_color_manual(name = "Model", values=col_models)+
      scale_fill_manual(name = "Model", values=col_models)

FigS6C<- forecastability %>%
  dplyr::filter(model == "Ensemble") %>%  ggplot(aes(x=omega, y=wis_scaled_relative_skill, col=season)) + 
  facet_grid(.~ disease) + 
  geom_point() +
   geom_smooth(aes(fill=season), method = "lm", se = TRUE)+
   theme_classic()+
    scale_color_manual(name = "Season", values=col_season)+
      scale_fill_manual(name = "Season", values=col_season)+
# theme(legend.position = "none")+
      xlab(expression(paste("Forecastability (",  Omega, ") (%)")))+
  ylab("Scaled Relative Skill")


FigS6 <-cowplot::plot_grid(FigS6A, FigS6B, FigS6C, ncol=1, labels="AUTO")

FigS6
```

### Fig S7. Coverage

```{r Fig6}
Fig6A<-forecastability %>% 
  dplyr::filter(model %in% c("Baseline", "Ensemble")) %>%
  ggplot(aes(x=omega, y=interval_coverage_50, col=model)) + 
  facet_grid(season~disease) + 
  geom_point() +
   geom_smooth(aes(fill=model), method = "lm", se = TRUE)+
   theme_classic()+
# theme(legend.position = "none")+
    xlab(expression(paste("Forecastability (",  Omega, ") (%)")))+
  ylab("Interval Coverage (50%)")+
    scale_color_manual(name = "Model", values=col_models)+
      scale_fill_manual(name = "Model", values=col_models)+
  geom_hline(yintercept = 0.5, linetype= "dashed")

Fig6B<-forecastability %>% 
  dplyr::filter(model %in% c("Baseline", "Ensemble")) %>%
  ggplot(aes(x=omega, y=interval_coverage_90, col=model)) + 
  facet_grid(season~disease) + 
  geom_point() +
   geom_smooth(aes(fill=model), method = "lm", se = TRUE)+
   theme_classic()+
# theme(legend.position = "none")+
    xlab(expression(paste("Forecastability (",  Omega, ") (%)")))+
  ylab("Interval Coverage (90%)")+
    scale_color_manual(name = "Model", values=col_models)+
      scale_fill_manual(name = "Model", values=col_models)+
  geom_hline(yintercept = 0.9, linetype= "dashed")
  

Fig6 <-cowplot::plot_grid(Fig6A, Fig6B, ncol=1, labels="AUTO")
Fig6
```






## Tables

### Table 1: Linear model results for forecast performance

```{r table1}
forecastability_nested <- forecastability %>%
  group_by(model, season, disease) %>%
  nest()

 lm_mae <- forecastability %>%
  group_by(model, season, disease) %>%
  nest() %>%
   mutate(data = map(data, ~ .x %>%
                    summarise(model = list(broom::tidy(lm(ae_median ~ omega)))))) %>% 
   unnest %>% 
   unnest
 
 lm_wis <- forecastability %>%
  group_by(model, season, disease) %>%
  nest() %>%
   mutate(data = map(data, ~ .x %>%
                    summarise(model = list(broom::tidy(lm(wis ~ omega)))))) %>% 
   unnest %>% 
   unnest %>%
   distinct() %>%
   arrange(disease, model, season)
 
  lm_skill <- forecastability %>%
  group_by(model, season, disease) %>%
  nest() %>%
   mutate(data = map(data, ~ .x %>%
                    summarise(model = list(broom::tidy(lm(wis_scaled_relative_skill ~ omega)))))) %>% 
   unnest %>% 
   unnest %>%
   distinct() %>%
   arrange(disease, model, season)
 
lm_results<- lm_mae %>% mutate(metric="MAE") %>% bind_rows(lm_wis %>% mutate(metric="WIS")) %>% filter(model %in% c("Baseline", "Ensemble")) %>% arrange(metric, disease, model, season) %>%
  mutate(across(where(is.numeric), signif, 3)) %>%
    filter(term=="omega") %>%
  dplyr::select(Metric=metric, Disease=disease, Model=model, Season=season, Estimate=estimate, `Standard Error`=std.error, Statistic=statistic, `p-value`=p.value)

lm_results_skill<-  lm_skill %>% mutate(metric="skill") %>% filter(model %in% c("Ensemble")) %>% arrange(metric, disease, model, season) %>%
  mutate(across(where(is.numeric), signif, 3)) %>%
    filter(term=="omega") %>%
  dplyr::select(Metric=metric, Disease=disease, Model=model, Season=season, Estimate=estimate, `Standard Error`=std.error, Statistic=statistic, `p-value`=p.value)


write_csv(lm_results, "results/lm_results_MMWR.csv")
write_csv(lm_results_skill, "results/lm_results_skill_MMWR.csv")

```


### Table 2: Linear model results for forecastability vs. cadence

```{r table2}
cadence_nested<- cadence %>% filter(season %notin% c("2022-2024"))  %>% 
  group_by(disease, cadence, season) %>%
  nest()

 lm_cadence <- cadence_nested%>%
   mutate(data = map(data, ~ .x %>%
                    summarise(model = list(broom::tidy(lm(omega ~ pop)))))) %>% 
   unnest %>% 
   unnest
 
write_csv(lm_cadence, "results/lm_cadence_MMWR.csv")

### Output for Table 2
lm_cadence %>% filter( term=="pop") %>%
  relocate(cadence, disease,  season) %>% arrange(cadence, disease,  season) %>%
  dplyr::select(!term) %>% 
  mutate_if(is.numeric, signif, 3)

```

### Table S1: ED Visits

```{r table-S1}
ed_nested <- ed_omega %>%
  group_by(season, disease) %>%
  nest()

 lm_ed <- ed_nested %>%
   mutate(data = map(data, ~ .x %>%
                    summarise(model = list(broom::tidy(lm(omega ~ log(pop))))))) %>% 
   unnest %>% 
   unnest
 
write_csv(lm_ed, "results/lm_ed_MMWR.csv")


lm_ed %>% mutate(metric="%ED visits") %>%
  filter(term=="log(pop)") %>%
  relocate(metric, disease, season) %>% arrange(metric, disease, season) %>%
  dplyr::select(!term) %>% 
  mutate_if(is.numeric, signif, 3)

```

### Table S2: Linear model results for forecast performance (July to June season definition)

```{r table1}
forecastability_nested <- forecastability_summer %>%
  group_by(model, season, disease) %>%
  nest()

 lm_mae <- forecastability_nested %>%
   mutate(data = map(data, ~ .x %>%
                    summarise(model = list(broom::tidy(lm(ae_median ~ omega)))))) %>% 
   unnest %>% 
   unnest
 
 lm_wis <- forecastability_nested %>%
   mutate(data = map(data, ~ .x %>%
                    summarise(model = list(broom::tidy(lm(wis ~ omega)))))) %>% 
   unnest %>% 
   unnest %>%
   distinct() %>%
   arrange(disease, model, season)
 
  lm_skill <- forecastability_nested %>%
   mutate(data = map(data, ~ .x %>%
                    summarise(model = list(broom::tidy(lm(wis_scaled_relative_skill ~ omega)))))) %>% 
   unnest %>% 
   unnest %>%
   distinct() %>%
   arrange(disease, model, season)
 
lm_results<- lm_mae %>% mutate(metric="MAE") %>% bind_rows(lm_wis %>% mutate(metric="WIS")) %>% filter(model %in% c("Baseline", "Ensemble")) %>% arrange(metric, disease, model, season) %>%
  mutate(across(where(is.numeric), signif, 3)) %>%
    filter(term=="omega") %>%
  dplyr::select(Metric=metric, Disease=disease, Model=model, Season=season, Estimate=estimate, `Standard Error`=std.error, Statistic=statistic, `p-value`=p.value)

lm_results_skill<-  lm_skill %>% mutate(metric="skill") %>% filter(model %in% c("Ensemble")) %>% arrange(metric, disease, model, season) %>%
  mutate(across(where(is.numeric), signif, 3)) %>%
    filter(term=="omega") %>%
  dplyr::select(Metric=metric, Disease=disease, Model=model, Season=season, Estimate=estimate, `Standard Error`=std.error, Statistic=statistic, `p-value`=p.value)


write_csv(lm_results, "results/lm_results.csv")
write_csv(lm_results_skill, "results/lm_results_skill.csv")

```

### Table S3: Horizon
```{r table-S3}
forecastability_horizon_nested <- forecastability_horizon %>%
  group_by(model, season, disease, horizon) %>%
  nest()


  lm_skill <- forecastability_horizon_nested %>%
   mutate(data = map(data, ~ .x %>%
                    summarise(model = list(broom::tidy(lm(wis_scaled_relative_skill ~ omega)))))) %>% 
   unnest %>% 
   unnest %>%
   distinct() %>%
   arrange(disease, model, season)
  
  lm_results_skill_horizon<-  lm_skill %>% mutate(metric="skill") %>% filter(model %in% c("Ensemble")) %>% arrange(metric, disease, model, season, horizon) %>%
  mutate(across(where(is.numeric), signif, 3)) %>%
    filter(term=="omega") %>%
  dplyr::select(Metric=metric, Disease=disease, Model=model, Season=season, Horizon=horizon, Estimate=estimate, `Standard Error`=std.error, Statistic=statistic, `p-value`=p.value)

lm_results_skill_horizon %>% mutate(Metric="Scaled Relative Skill") %>%
  filter(Model %in% c("Ensemble", "Baseline")) %>%
  relocate(Metric, Disease, Model, Season, Horizon) %>% arrange(Metric, Disease, Model, Season, Horizon) %>%
  mutate_if(is.numeric, signif, 3)
  
write_csv(lm_results_skill_horizon, "results/lm_results_skill_horizon_MMWR.csv")


```


### Table S4: Coverage

```{r table-S4}
forecastability_nested <- forecastability %>%
  group_by(model, season, disease) %>%
  nest()

 lm_coverage50 <- forecastability %>%
  group_by(model, season, disease) %>%
  nest() %>%
   mutate(data = map(data, ~ .x %>%
                    summarise(model = list(broom::tidy(lm(interval_coverage_50 ~ omega)))))) %>% 
   unnest %>% 
   unnest
 
 lm_coverage90 <- forecastability %>%
  group_by(model, season, disease) %>%
  nest() %>%
   mutate(data = map(data, ~ .x %>%
                    summarise(model = list(broom::tidy(lm(interval_coverage_90 ~ omega)))))) %>% 
   unnest %>% 
   unnest
 

write_csv(lm_coverage50, "results/lm_coverage50_MMWR.csv")
write_csv(lm_coverage90, "results/lm_coverage90_MMWR.csv")

lm_coverage50 %>% mutate(metric="Interval Coverage (50%)") %>% bind_rows(lm_coverage90 %>% mutate(metric="Interval Coverage (90%)")) %>% 
  filter(model %in% c("Ensemble", "Baseline"), term=="omega") %>%
  relocate(metric, disease, model, season) %>% arrange(metric, disease, model, season) %>%
  dplyr::select(!term) %>% 
  mutate_if(is.numeric, signif, 3)

```



## Session Info
```{r session-info}
sessionInfo()
```