---
title: "Generate Figures and Tables"
author: "Lauren White (lauren.white@cdph.ca.gov)"
date: '2024-12-05'
output:
  html_document:
    toc: true
    toc_float: true
---

Code to reproduce figure and table results in manuscripts. Can be run sequentially after `load_data.R`, `calculate_forecastability.Rmd`, and `forecast_eval.Rmd` or as a standalone using output already saved locally to the `/results` folder for convenience.

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggrepel)
library(dplyr)
library(MetBrewer)
library(readr)
library(cowplot)
library(tidyr)
library(purrr)
library(segmented)
library(astsa)
library(ForeCA)
library(forecast)

`%notin%` <- Negate(`%in%`)
```

### Settings for `foreCA` `Omega` function
```{r}
# spectrum control
sc <- list(method = "mvspec")

# entropy control
ec <- list(prior.weight = 1e-2)

confidence<-0.05
```

### Load forecastability data frame
(if not already loaded in environment from running other Rmd scripts)
```{r}
if(!exists("forecastability")) {
  forecastability<-read_csv("results/forecastability_results.csv")
}

if(!exists("combo")) {
  combo<-read_csv("results/combo_omega.csv")
}

if(!exists("hcai_combined")) {
  hcai_combined<-read_csv("results/hcai_combined.csv")
}

if(!exists("hcai_state_omega")) {
  hcai_state_omega<-read_csv("results/hcai_state_omega.csv")
}

if(!exists("hcai_burden")) {
  hcai_burden<-read_csv("results/hcai_burden.csv")
}

if(!exists("state_covid_cadence")) {
  state_covid_cadence<-read_csv("results/state_covid_cadence.csv")
}

if(!exists("state_flu_cadence")) {
  state_flu_cadence<-read_csv("results/state_flu_cadence.csv")
}

if(!exists("flu_subtypes")) {
flu_subtypes<-read_csv("data/CDPH_flu_subtypes.csv") %>% rename(season=Season)
}

if(!exists("covid_hosp_full")) {
covid_hosp_full<-read_csv("data/covid_hosp.csv") 
}

if(!exists("flu_hosp_full")) {
flu_hosp_full<-read_csv("data/flu_hosp.csv") 
}
```


### Predefine color palettes
```{r color-palettes}

col_diseases<-met.brewer("Hiroshige",2)

col_models<-met.brewer("Cassatt1",2)

col_season<-met.brewer("Hokusai1",6)
names(col_season)<-c("2017-2018", "2018-2019", "2019-2020", "2021-2022", "2022-2023", "2023-2024") 

col_subtype<-met.brewer("Isfahan2", 4)
                     
```


## Figure 1
https://stackoverflow.com/questions/61761976/plotting-multiple-acf-with-ggplot
```{r fig1}
set.seed(1234)
flu_ts<-flu_hosp_full%>% filter(geo_value=="us") %>% 
  arrange(time_value) %>%
  mutate(days= time_value-as.Date("2022-02-02")) 
flu_ts_state<-flu_hosp_full %>% filter(geo_value=="md") %>% 
  arrange(time_value) %>%
  mutate(days= time_value-as.Date("2022-02-02")) 

flu_ts %>% 
    pull(weekly_hosp) %>% ndiffs(test="kpss", alpha=confidence)
flu_ts %>% 
    pull(weekly_hosp) %>% ndiffs(test="adf", alpha=confidence)

flu_ts_state%>% 
    pull(weekly_hosp) %>% ndiffs(test="adf", alpha=confidence)
flu_ts_state%>% 
    pull(weekly_hosp) %>% ndiffs(test="kpss", alpha=confidence)

covid_ts<-covid_hosp_full %>% filter(geo_value=="us") %>% mutate(days= time_value-as.Date("2022-02-02")) %>% filter(days>=0)

flu_ts<- flu_ts %>% mutate(weekly_hosp_lag= weekly_hosp-lag(weekly_hosp)) %>% 
  mutate(weekly_hosp_lag2= weekly_hosp_lag-lag(weekly_hosp_lag)) %>% 
  drop_na()


flu_ts_state<-flu_ts_state %>% mutate(weekly_hosp_lag= weekly_hosp-lag(weekly_hosp)) %>%
  mutate(weekly_hosp_lag2= weekly_hosp_lag-lag(weekly_hosp_lag)) %>% 
  drop_na()

fore_flu<- as.numeric(Omega(flu_ts$weekly_hosp_lag, spectrum.control = sc, entropy.control = ec))
fore_flu_state<- as.numeric(Omega(flu_ts_state$weekly_hosp_lag, spectrum.control = sc, entropy.control = ec))

fore_covid<- combo %>% filter(season=="2022-2024", disease=="COVID-19", location=="us") %>% pull(omega)

#Random ts
days<-seq(1,nrow(flu_ts), by=1)
y<-rnorm(length(days),mean=1,sd=10)
random_ts<- data.frame(days,y)
random_ts %>%     
  pull(y) %>% ndiffs(test="kpss", alpha=confidence)
random_ts %>%     
  pull(y) %>% ndiffs(test="adf", alpha=confidence)
random_ts<- random_ts %>% mutate(lag_y= y-lag(y)) %>% 
  mutate(lay_y=lag_y-lag(lag_y)) %>% drop_na()


fore_rand<-as.numeric(Omega(random_ts$lag_y, spectrum.control = sc, entropy.control = ec))


Fig1A<- random_ts %>% ggplot(aes(days,y)) + geom_line() + theme_classic()+ ggtitle("Signal") +xlab("")+ylab("")
rand_acf<- data.frame(lag=acf(random_ts$y, main="", lag.max=365)$lag, acf=acf(random_ts$y, main="",lag.max=365)$acf) %>% filter(lag%%14 == 0)
Fig1B<- rand_acf %>% ggplot(aes(x=lag, y=acf)) +
  geom_segment(mapping = aes(xend = lag, yend = 0)) +
          geom_point()+ theme_classic() + ggtitle("Autocorrelation") +xlab("")+ylab("")
rand_spec<-data.frame(freq=mvspec(random_ts$y, log="no")$freq, spec= mvspec(random_ts$y, log="no")$spec) %>% 
  filter(freq<=0.025)

Fig1C<- rand_spec %>% ggplot(aes(x=freq, y=spec)) + geom_line() + theme_classic()+
  annotate(geom="text", x=0.02, y=250, label=paste(round(fore_rand,2),"%"),
              color="red")+
  ggtitle("Spectral Density") +xlab("")+ylab("")

# cowplot::plot_grid(A, B, C, ncol=1, label="AUTO")


#Flu state ts
Fig1D<- flu_ts_state %>% ggplot(aes(days, weekly_admits_p100K)) + geom_line()+ theme_classic() +xlab("")+ylab("")
#acf(flu_ts_state$weekly_admits_p100K, main="")
covid_acf<- data.frame(lag=acf(flu_ts_state$weekly_admits_p100K, lag.max=365)$lag, acf=acf(flu_ts_state$weekly_admits_p100K, lag.max=365)$acf) %>% filter(lag%%14 == 0)
Fig1E<- covid_acf %>% ggplot(aes(x=lag, y=acf)) +
  geom_segment(mapping = aes(xend = lag, yend = 0)) +
          geom_point()+ theme_classic() +xlab("")+ylab("")
flu_state_spec<- data.frame(freq=mvspec(flu_ts_state$weekly_admits_p100K, log="no")$freq, spec= mvspec(flu_ts_state$weekly_admits_p100K, log="no")$spec) %>% 
  filter(freq<=0.025)
Fig1F<- flu_state_spec %>% ggplot(aes(x=freq, y=spec)) + geom_line() + theme_classic()+
  annotate(geom="text", x=0.02, y=200, label=paste(round(fore_flu_state,2),"%"),
              color="red") +
  xlab("")+ylab("")


#Flu ts  
Fig1G<- flu_ts %>% ggplot(aes(days, weekly_admits_p100K)) + geom_line()+ theme_classic() +xlab("")+ylab("")
#acf(flu_ts$weekly_admits_p100K, main="")
flu_acf<- data.frame(lag=acf(flu_ts$weekly_admits_p100K, lag.max=365)$lag, acf=acf(flu_ts$weekly_admits_p100K, lag.max=365)$acf) %>% filter(lag%%14 == 0)
Fig1H<- flu_acf %>% ggplot(aes(x=lag, y=acf)) +
  geom_segment(mapping = aes(xend = lag, yend = 0)) +
          geom_point()+ theme_classic() +xlab("")+ylab("")
flu_spec<- data.frame(freq=mvspec(flu_ts$weekly_admits_p100K, log="no")$freq, spec= mvspec(flu_ts$weekly_admits_p100K, log="no")$spec) %>% 
  filter(freq<=0.025)
Fig1I<- flu_spec %>% ggplot(aes(x=freq, y=spec)) + geom_line() + theme_classic()+
  annotate(geom="text", x=0.02, y=300, label=paste(round(fore_flu,2),"%"),
              color="red") + 
  xlab("")+ylab("")

#Sine ts
sine_ts<- data.frame(days, y= sin(0.05*days))
sine_ts %>%     
  pull(y) %>% ndiffs(test="kpss", alpha=confidence)
sine_ts %>%     
  pull(y) %>% ndiffs(test="adf", alpha=confidence)
sine_ts<- sine_ts %>% mutate(lag_y= y-lag(y)) %>%
   mutate(lag_y= lag_y-lag(lag_y)) %>% drop_na()

fore_sine<- as.numeric(Omega(sine_ts$lag_y, spectrum.control = sc, entropy.control = ec))

Fig1J<- sine_ts %>% ggplot(aes(days,y)) + geom_line() + theme_classic() +xlab("Days")+ylab("")
acf(sine_ts$y, main="")
mvspec(sine_ts$y, log="no")
sine_acf<- data.frame(lag=acf(sine_ts$y, lag.max=365)$lag, acf=acf(sine_ts$y, lag.max=365)$acf) %>% filter(lag%%14 == 0)

Fig1K<- sine_acf %>% ggplot(aes(x=lag, y=acf)) +
  geom_segment(mapping = aes(xend = lag, yend = 0)) +
          geom_point()+ theme_classic() + xlab("Lag (Days)") + ylab("")
sine_spec<- data.frame(freq=mvspec(sine_ts$y, log="no")$freq, spec= mvspec(sine_ts$y, log="no")$spec) %>% 
  filter(freq<=0.025)
Fig1L<- sine_spec %>% ggplot(aes(x=freq, y=spec)) + geom_line() + theme_classic()+
  annotate(geom="text", x=0.02, y=60, label=paste(round(fore_sine,2),"%"),
              color="red")+
  xlab("Frequency")+ylab("")

cowplot::plot_grid(Fig1A, Fig1B, Fig1C, Fig1D, Fig1E, Fig1F, Fig1G, Fig1H, Fig1I, Fig1J, Fig1K, Fig1L, ncol=3,labels= "AUTO") # labels=c("A Signal", "B Autocorrelation", "C Spectral Density", "D", "E", "F", "G", "H", "I", "J", "K", "L"))


```



## Figure 2

### Plot figures

```{r fig2}
Fig2A<- hcai_combined %>% filter(season %in% c("2017-2018", "2018-2019", "2019-2020", "2021-2022")) %>% ggplot(aes(x=log(pop), y=omega, col=season)) +
  geom_point()+ 
  theme_classic()+  
  # geom_text_repel(aes(label=location))+ 
  # ggtitle("HCAI Weekly Influenza Admissions Time Series Forecastability (County and State)")+
  xlab("Natural logarithm of population size")+
  ylab(expression(paste("Forecastability (",  Omega, ") (%)")))+
  geom_smooth(aes(fill=season), method="gam", formula= y~s(x, bs = "cs"))+
  scale_color_manual(name = "Season", values=col_season)+
  scale_fill_manual(name = "Season", values=col_season)+
  theme(legend.position="bottom")+
  # xlim(c(9,20))+
  ylim(c(0,45))


Fig2B<- hcai_state_omega %>% mutate(season_dummy=season) %>% 
  separate(season_dummy, into = c("season_start", "season_end"), sep = "-") %>% 
  left_join(hcai_burden, by=c("location", "season")) %>% 
  left_join(flu_subtypes, by="season")%>%
  filter(location=="California", season %notin% c("1999-2000", "2020-2021", "2022-2023")) %>% ggplot(aes(x=log(max_admits), y=omega)) + 
  geom_smooth(method = "lm", se = TRUE, alpha=0.3)+
  geom_point(aes(col=CDPH), size=2)+ 
  geom_text_repel(aes(label=season_start, col=CDPH))+ 
  theme_classic()+
  xlab("Natural logarithm of peak weekly admissions")+
  ylab(expression(paste("Forecastability (",  Omega, ") (%)")))+
  scale_color_manual(name = "Predominant Influenza Subtype", values=col_subtype)+
  theme(legend.position="bottom")+
  ylim(c(0,45))

cowplot::plot_grid(Fig2A, Fig2B, ncol=1, labels="AUTO", align="h")
```

## Figure 4: Sampling Frequency
```{r fig4}
cadence<- state_flu_cadence %>% bind_rows(state_covid_cadence)

Fig4<- cadence %>% filter(season %notin% c("2022-2024")) %>% 
  ggplot(aes(x=log(pop), y=omega, col=season)) +
  geom_point()+ 
  facet_grid( disease ~cadence)+
  theme_classic()+  
  geom_smooth(aes(fill=season), method = "lm", se = TRUE, alpha=0.3)+
  geom_text_repel(aes(label=state))+ 
  scale_color_manual(name = "Season", values=col_season)+
  scale_fill_manual(name = "Season", values=col_season)+
  ylim(c(0,45))+
  theme(legend.position="bottom")+
  xlab("Natural logarithm of population size")+
  ylab(expression(paste("Forecastability (",  Omega, ") (%)")))

Fig4


# SupFig2<-state_covid_cadence %>% 
#   filter(season != "2022-2024")%>%
#   ggplot(aes(x=log(pop), y=omega, col=season)) +
#   facet_wrap(data_type ~.)+
#   geom_point()+ 
#   theme_classic()+  
#   geom_smooth(aes(fill=season), method = "lm", se = TRUE, alpha=0.3)+
#   geom_text_repel(aes(label=state))+ 
#   scale_color_manual(name = "Season", values=col_season)+
#   scale_fill_manual(name = "Season", values=col_season)+
#   ylim(c(0,45))+
#   theme(legend.position="bottom")+
#   xlab("Natural logarithm of population size")+
#   ylab(expression(paste("Forecastability (",  Omega, ") (%)")))


```


## Supplemental Figure 1

```{r SupFig1}
SupFig1<-hcai_state_omega %>% mutate(season_dummy=season) %>% 
  separate(season_dummy, into = c("season_start", "season_end"), sep = "-") %>%  
  left_join(hcai_burden, by=c("location", "season")) %>% 
    left_join(flu_subtypes, by="season")%>%
  filter(location=="California", season %notin% c("1999-2000", "2020-2021", "2022-2023")) %>% ggplot(aes(x=log(total_admits), y=omega)) + 
  geom_smooth(method = "lm", se = TRUE, alpha=0.3)+
  geom_text_repel(aes(label=season_start, col=CDPH))+ 
  geom_point(aes(col=CDPH), size=2)+ 
  theme_classic()+
  xlab("Natural logarithm of cumulative hospital admissions")+
  ylab(expression(paste("Forecastability (",  Omega, ") (%)")))+
  scale_color_manual(name = "Predominant Influenza Subtype", values=col_subtype)

SupFig1


```


## Figure 3
```{r fig3}

Fig3A<-combo %>% filter(season %in% c("2022-2023", "2023-2024"), location_level=="state") %>%
  ggplot(aes(x=log_pop, y=omega, col=season))  +
  geom_point(alpha=0.5) +
  facet_grid(.~disease) +
  # geom_line(data = flu_piece, color = col_diseases[2])+
  # geom_line(data = covid_piece, color = col_diseases[1])+
  geom_smooth(aes(fill=season, col=season), method="gam", formula= y~s(x, bs = "cs"))+

  xlab("Natural logarithm of population size")+
  ylab(expression(paste("Forecastability (",  Omega, ") (%)")))+
    scale_color_manual(name = "Season", values=col_season)+
      scale_fill_manual(name = "Season", values=col_season)+
  theme_classic()

Fig3B<-forecastability %>% 
  dplyr::filter(model %in% c("Baseline", "Ensemble")) %>%
  ggplot(aes(x=omega, y=ae_median, col=model)) + 
  facet_grid(season~disease) + 
  geom_point() +
   geom_smooth(aes(fill=model), method = "lm", se = TRUE)+
   theme_classic()+
# theme(legend.position = "none")+
    xlab(expression(paste("Forecastability (",  Omega, ") (%)")))+
  ylab("MAE on Log Scale")+
    scale_color_manual(name = "Model", values=col_models)+
      scale_fill_manual(name = "Model", values=col_models)


Fig3C<-forecastability %>%
  dplyr::filter(model %in% c("Baseline", "Ensemble")) %>%  ggplot(aes(x=omega, y=interval_score, col=model)) + facet_grid(season~disease) + 
  geom_point() +
   geom_smooth(aes(fill=model), method = "lm", se = TRUE)+
   theme_classic()+
# theme(legend.position = "none")+
      xlab(expression(paste("Forecastability (",  Omega, ") (%)")))+
  ylab("Mean WIS on Log Scale")+
    scale_color_manual(name = "Model", values=col_models)+
      scale_fill_manual(name = "Model", values=col_models)

Fig3D<- forecastability %>%
  dplyr::filter(model == "Ensemble") %>%  ggplot(aes(x=omega, y=scaled_rel_skill, col=season)) + 
  facet_grid(.~ disease) + 
  geom_point() +
   geom_smooth(aes(fill=season), method = "lm", se = TRUE)+
   theme_classic()+
    scale_color_manual(name = "Season", values=col_season)+
      scale_fill_manual(name = "Season", values=col_season)+
# theme(legend.position = "none")+
      xlab(expression(paste("Forecastability (",  Omega, ") (%)")))+
  ylab("Scaled Relative Skill")


Fig3 <-cowplot::plot_grid(Fig3A, Fig3B, Fig3C, Fig3D, ncol=1, labels="AUTO")

Fig3
```

## Supplementary Figure 2
```{r sup-fig2}


SupFig2A<-forecastability %>% 
  dplyr::filter(model %notin% c("MOBS-GLEAM", "CEPH-Rtrend")) %>%
  ggplot(aes(x=log(pop), y=ae_median, col=model))+ facet_grid(season~disease)+ geom_point()+
   theme_classic()+
   geom_smooth(method = "lm", se = TRUE)+
theme(legend.position = "none")+
  xlab("ln(Population size)")+
  ylab("Mean Absolute Error")


SupFig2B<-forecastability %>% 
  dplyr::filter(model %notin% c("MOBS-GLEAM", "CEPH-Rtrend")) %>%
  ggplot(aes(x=log(pop), y=interval_score, col=model))+ facet_grid(season~disease)+ geom_point()+
   theme_classic()+
   geom_smooth(method = "lm", se = TRUE)+
    xlab("ln(Population size)")+
theme(legend.position = "none")+
  ylab("Mean Weighted Interval Score")


legend_plot<- forecastability %>% 
    dplyr::filter(model %notin% c("MOBS-GLEAM", "CEPH-Rtrend")) %>%
  ggplot(aes(x=log(pop), y=ae_median, col=model))+ facet_wrap(season~disease)+ geom_point()+
   theme_classic()+
   geom_smooth(method = "lm", se = TRUE)+
    xlab("ln(Population size)")+
  ylab("Median Absolute Error")+
  theme(legend.position="bottom")

SupFig2_legend <- get_plot_component(legend_plot, "guide-box", return_all = TRUE)

preSupFig2<-cowplot::plot_grid(SupFig2A, SupFig2B, nrow=1, labels="AUTO")
SupFig2<-cowplot::plot_grid(preSupFig2, SupFig2_legend[[3]], nrow=2, rel_heights = c(2,0.5))


SupFig2
```


## Table 1: Linear model results

```{r table1}
forecastability_nested <- forecastability %>%
  group_by(model, season, disease) %>%
  nest()

 lm_mae <- forecastability %>%
  group_by(model, season, disease) %>%
  nest() %>%
   mutate(data = map(data, ~ .x %>%
                    summarise(model = list(broom::tidy(lm(ae_median ~ omega)))))) %>% 
   unnest %>% 
   unnest
 
 lm_wis <- forecastability %>%
  group_by(model, season, disease) %>%
  nest() %>%
   mutate(data = map(data, ~ .x %>%
                    summarise(model = list(broom::tidy(lm(interval_score ~ omega)))))) %>% 
   unnest %>% 
   unnest %>%
   distinct() %>%
   arrange(disease, model, season)
 
  lm_skill <- forecastability %>%
  group_by(model, season, disease) %>%
  nest() %>%
   mutate(data = map(data, ~ .x %>%
                    summarise(model = list(broom::tidy(lm(scaled_rel_skill ~ omega)))))) %>% 
   unnest %>% 
   unnest %>%
   distinct() %>%
   arrange(disease, model, season)
 
lm_results<- lm_mae %>% mutate(metric="MAE") %>% bind_rows(lm_wis %>% mutate(metric="WIS")) %>% filter(model %in% c("Baseline", "Ensemble")) %>% arrange(metric, disease, model, season) %>%
  mutate(across(where(is.numeric), signif, 3)) %>%
    filter(term=="omega") %>%
  dplyr::select(Metric=metric, Disease=disease, Model=model, Season=season, Estimate=estimate, `Standard Error`=std.error, Statistic=statistic, `p-value`=p.value)

lm_results_skill<-  lm_skill %>% mutate(metric="skill") %>% filter(model %in% c("Ensemble")) %>% arrange(metric, disease, model, season) %>%
  mutate(across(where(is.numeric), signif, 3)) %>%
    filter(term=="omega") %>%
  dplyr::select(Metric=metric, Disease=disease, Model=model, Season=season, Estimate=estimate, `Standard Error`=std.error, Statistic=statistic, `p-value`=p.value)


write_csv(lm_results, "results/lm_results_MMWR.csv")
write_csv(lm_results_skill, "results/lm_results_skill__MMWR.csv")

```

## Session Info
```{r session-info}
sessionInfo()
```